#  Multivariate analyses

First load the required packages and data. Note:, from the univariate analyses the analysis data set for the lab parameters has been updated to include transformed variables. Therefore, we load the second iteration of the data `data/ADLB_02.rds`. 

```{r, echo = FALSE, message = FALSE, warning = FALSE }
library(tidyr)
library(dplyr)
library(ggplot2)
library(purrr)
library(tidyselect)
library(corrr) ## tidy correlation
library(patchwork)
library(ggpubr)
library(ggrepel)
library(here)
library(dendextend)
library(gt)
library(gtExtras)

## for network plot
options(ggrepel.max.overlaps = 100)

source(here("R", "fun_compare_dist_plot.R"))
source(here("R", "fun_compare_mult_dists.R"))
source(here("R", "fun_ida_trans.R"))

## Load the datasets - make sure to load the correct ADLB2 
## TODO: have a make workflow to ensure dependencies run in correct order

ADSL <- readRDS(here::here("data","IDA", "ADSL_01.rds"))
ADLB <- readRDS(here::here("data","IDA", "ADLB_02.rds"))

```



## V1: Association with structural variables

Attached the required structural variables to the lab data. 

```{r mvi01, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}

## Join with ADSL for structural variables 
ADSL01 <-
  ADSL |> 
  select(USUBJID, AGEGR01C, AGE, SEXC, SEX) 

dat <- ADLB |> 
  left_join(ADSL01, by = "USUBJID")
```



A scatterplot of each predictor with age, with different panels for males and females have been constructed. Associated Spearman correlation coefficients have been computed.

The remaining predicotrs are reported in appendix @sec-V1-appendix.

### Key predictors


```{r mvi03, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}
#| layout-ncol: 3 

## plots for key predictors - remove age to avoid age vs age plot
key_predictors <- dat |> 
  filter(!PARAMCD == "AGE") |>
  filter(KEY_PRED_FL02 == "Y") |>
  group_by(PARAMCD) |>
  group_map(~ plot_assoc_by(.x), .keep = TRUE)

## print out. TODO: ask GH which plots to save
for (plts in key_predictors) {
  print(plts)
}

```


### Predictors of medium importance

```{r mvi04, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}
#| layout-ncol: 3 

medium_predictors <- dat |> 
  filter(MED_PRED_FL02 == "Y") |>
  group_by(PARAMCD) |>
  group_map(~ plot_assoc_by(.x), .keep = TRUE)

for (plts in medium_predictors) {
  print(plts)
}

```


## V2: Correlation coefficients between all predictors

Calculate correlation matrix using Spearman correlation coefficient. 

```{r mcorrs, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}

## tidy parameters in to a wide data set
dat_wide <- dat |>
  filter(KEY_PRED_FL02 == "Y" | MED_PRED_FL02 == "Y" | REM_PRED_FL02 == "Y") |>
  select(USUBJID, PARAMCD, AVAL, SEX) |>
  pivot_wider(names_from = PARAMCD, values_from = AVAL, values_fill = NA) 

## calculate corr matrix using spearman cc
corr_spearman <- 
  dat_wide |> 
  select(-USUBJID) |>
  correlate(use = "pairwise.complete.obs",
            method = "spearman",
            diagonal = NA)


```

The Spearman correlation coefficients are depicted in a quadratic heat map:

```{r plotcorr, message=FALSE, warning=FALSE, echo=FALSE, fig.width = 12, fig.height=12}


corr_spearman |> rearrange(method = "HC") |>
  autoplot(
#         method = "PCA",
         triangular = "lower",
         barheight = 20) +
  theme(
    axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust = 1, size = 8),
    axis.text.y = ggplot2::element_text(vjust = 1, hjust = 1, size = 9)
  )

```



```{r}
## If we want to write to file or save
# spearman_corr |>
#   shave(upper = TRUE) |>
#   fashion() |>
#   gt::gt()
```



### VE1: Comparing nonparametric and parametric predictor correlation

```{r corr_setup, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 12, fig.height=12}
# differences of pearson and spearman correlations to check for outliers

## cast as a matrix 
corr_spearman_mat <- 
  corr_spearman |>  
  as_matrix()

## calculate pearson cc
corr_pearson <- 
  dat_wide |> 
  select(-USUBJID) |>
  correlate(use = "pairwise.complete.obs",
            method = "pearson",
            diagonal = NA)

## cast as matrix 
corr_pearson_mat <- 
  corr_pearson |> 
  as_matrix()


## calculate the difference 
corrd <- corr_spearman_mat - corr_pearson_mat

## cast as tidy data frame TODO: could do this through a join?
corr_difference <- corrd |> as_cordf()
```


Plot the difference matrix 

```{r plotcorr_diff, message=FALSE, warning=FALSE, echo=FALSE, fig.width = 12, fig.height=12}

## plot

corr_difference |> rearrange(method = "HC") |>
  autoplot(
#         method = "PCA",
         triangular = "lower",
         barheight = 20) +
  theme(
    axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust = 1, size = 8),
    axis.text.y = ggplot2::element_text(vjust = 1, hjust = 1, size = 9)
  )

```

Plot the matrix but suppress differences less then 0.1. 

```{r plotcorr02, message=FALSE, warning=FALSE, echo=FALSE, fig.width = 12, fig.height=12}

# sparsified differences of correlation coefficients
corr_difference_conditional <-
  corr_difference |> 
  stretch() |>
  mutate(r = if_else(abs(r) > 0.1, r, 0)) |>
  retract() |>
  as_cordf()

## todo: remove zero entries?
corr_difference_conditional |> 
  rearrange(method = "HC") |>
  autoplot(
#         method = "PCA",
         triangular = "lower",
         barheight = 20) +
  theme(
    axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust = 1, size = 8),
    axis.text.y = ggplot2::element_text(vjust = 1, hjust = 1, size = 9)
  )


```


```{r plotcorr03, message=FALSE, warning=FALSE, echo=FALSE, fig.width = 12, fig.height=12}

#corr_difference_conditional |>
#  shave() |>
#  rplot(print_cor = TRUE)


#corrd_sp <- corrd
#corrd_sp[abs(corrd)<0.1] <-0

#ggcorrplot(corrd_sp, tl.cex=5, tl.srt=90)

#ggcorrplot(corrd_sp, type = "lower", outline.col = "white")

```


Predictor pairs for which Spearman and Pearson correlation coefficients differ by more than 0.1 correlation units will be depicted in scatterplots. First report the table of predictor parirs where `r` is greater than 0.1. 

```{r pred_corr, message=FALSE, warning=FALSE, echo=FALSE}

## COMMENT @Mark can we print pearson and spearman correlation coefficients into/over the graphs?

combos <-
  corr_difference_conditional |>
  shave() |>
  stretch() |>
  filter(r > 0.1) 

combos |>
  gt::gt() |>
     gt::fmt_number(
      columns = c(r),
      decimals = 2
    ) |> 
  gtExtras::gt_theme_538()

```

Then plot the scatter plots. 

```{r ve3, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}
#| layout-ncol: 3 

## Data prep

combos01 <- combos |>
  mutate(grp = row_number())

combos_spearman <-
  corr_spearman |>
  stretch() |>
  rename(spearman = r)

combos_pearson <-
  corr_pearson |>
  stretch() |>
  rename(pearson = r)


combos02 <- combos01 |>
  left_join(combos_spearman) |>
  left_join(combos_pearson)
```



```{r ve3_02, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}
#| layout-ncol: 3 

## Plot the data. 

source(here("R", "V2-mv-descriptions.R"))

combos3 <-
  combos02 |>
  group_by(grp) |>
  group_map(~ plot_scatter_corr(.x), .keep = TRUE)

```


```{r ve3_03, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}
#| layout-ncol: 3 

## print out. TODO: ask GH which plots to save
for (plts in combos3) {
  print(plts)
}
```



### VE2: Variable clustering

A variable clustering analysis has been performed to evaluate which predictors are closely associated. The dendrogram groups predictors by their correlation.

```{r varclust, message=FALSE, warning=FALSE, echo=FALSE, fig.width = 12, fig.height=12}
source(here("R", "VE2-varclust.R"))

ve2_varclust(corr_spearman)

#vc_bact<-Hmisc::varclus(as.matrix(dat))
#plot(vc_bact, cex=0.7, hang=0.01)
```


This could also be displayed as a network plot

```{r varclust02, message=FALSE, warning=FALSE, echo=FALSE, fig.width = 12, fig.height=6}
#options(ggrepel.max.overlaps = 100)
corr_spearman |> network_plot()
```


```{r varclust03, message=FALSE, warning=FALSE, echo=FALSE, fig.width = 12, fig.height=6}
# plot network with min cor of 0.3. 
# corr_spearman |> network_plot(min_cor = 0.3)
```



In the following scatterplots we show predictor pairs with Spearman correlation coefficients greater than 0.8:

```{r varclust04, message=FALSE, warning=FALSE, echo=FALSE, fig.width = 12, fig.height=6}

combos_spearman_02 <-
  corr_spearman |>
  shave() |>
  stretch() |>
  rename(spearman = r) |>
  filter(abs(spearman) > 0.8) |>
  mutate(grp = row_number())

```

```{r varclust05, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}
#| layout-ncol: 4 

source(here("R", "V2-mv-descriptions.R"))

combos_spearman_03 <-
  combos_spearman_02 |>
  group_by(grp) |>
  group_map(~ plot_scatter_corr(.x, pearson_flag = FALSE), .keep = TRUE)


## print out. TODO: ask GH which plots to save
for (plts in combos_spearman_03 ) {
  print(plts)
}


```



### VE3: Redundancy

Variance inflation factors (VIF) will be computed between the candidate predictors. This will be done for the three possible candidate models, and using all complete cases in the respective candidate predictor sets. Since $VIF = (1-R^2)^{-1}$, we also report the multiple R-squared values. 

Redundancy was further explored by computing parametric additive models for each predictor in the key predictor model and the extended predictor model. VIFs and multiple $R^2$ are reported from those models, again for the three predictor sets.

Note, the all predictor model is reported in appendix @sec-VE3-appendix. 

#### VIF for key predictor model

```{r vif, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}


dat_wide <- dat |>
  filter(KEY_PRED_FL02 == "Y") |>
  select(USUBJID, PARAMCD, AVAL, SEX) |>
  pivot_wider(names_from = PARAMCD, values_from = AVAL, values_fill = NA) 

dat_vif <- dat_wide |>
  select(-USUBJID)

formula <- as.formula(paste(c("~",paste(names(dat_vif), collapse="+")), collapse=""))
 
#formula

red <- Hmisc::redun(formula, data = dat_vif, nk=0, pr=FALSE)
vif <- 1/(1-red$rsq1)
 
#cat("\nAvailable samvple size:\n", red$n, " (", #round(100*red$n/nrow(dat),2), "%)\n")

vif_df <- as_tibble(vif, rownames = "PARAMCD") |>
  rename(vif = value)

red_df <- as_tibble(red$rsq1, rownames = "PARAMCD") |>
  rename(r2 = value)

tab_df <- vif_df |> left_join(red_df, by = "PARAMCD")


```

The available sample size is `r red$n` (`r round(100*red$n/nrow(dat),2)` %).

```{r vif02, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}

#vif_df |> gt::gt(caption = "Variance inflation factors")
#red_df |> gt::gt(caption = "Multiple R-squared")

tab_df |> 
  gt() |>
    gt::cols_label(
    PARAMCD = "Parameter code",
    r2 = "Multiple R-squared",
    vif = "Variance inflation factor"
  ) |>
  fmt_number(
    columns = c(r2),
    decimals = 2
  ) |>
  fmt_number(
    columns = c(vif),
    decimals = 1
  ) |>
  gt_theme_538()
  

```


#### VIF for model with key predictors and predictors of medium importance

```{r vif03, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}

dat_wide <- dat |>
  filter(KEY_PRED_FL02 == "Y" | MED_PRED_FL02 == "Y") |>
  select(USUBJID, PARAMCD, AVAL, SEX) |>
  pivot_wider(names_from = PARAMCD, values_from = AVAL, values_fill = NA) 


dat_vif <- dat_wide |>
  select(-USUBJID)

formula <- as.formula(paste(c("~",paste(names(dat_vif), collapse="+")), collapse=""))
 
#formula

red <- Hmisc::redun(formula, data = dat_vif, nk=0, pr=FALSE)
vif <- 1/(1-red$rsq1)
 
#cat("\nAvailable samvple size:\n", red$n, " (", #round(100*red$n/nrow(dat),2), "%)\n")

vif_df <- as_tibble(vif, rownames = "PARAMCD") |>
  rename(vif = value)

red_df <- as_tibble(red$rsq1, rownames = "PARAMCD") |>
  rename(r2 = value)

tab_df <- vif_df |> left_join(red_df, by = "PARAMCD")


```

The available sample size is `r red$n` (`r round(100*red$n/nrow(dat),2)` %).

```{r vif04, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}

#vif_df |> gt::gt(caption = "Variance inflation factors")
#red_df |> gt::gt(caption = "Multiple R-squared")

tab_df |> 
  gt() |>
    gt::cols_label(
    PARAMCD = "Parameter code",
    r2 = "Multiple R-squared",
    vif = "Variance inflation factor"
  ) |>
  fmt_number(
    columns = c(r2),
    decimals = 2
  ) |>
  fmt_number(
    columns = c(vif),
    decimals = 1
  ) |>
  gt_theme_538()
  

```

#### VIF for all predictor model

See appendix @sec-VE3-appendix. 


#### Redundancy by parametric additive model: key predictor model

```{r key_vif, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}


dat_wide <- dat |>
  filter(KEY_PRED_FL02 == "Y") |>
  select(USUBJID, PARAMCD, AVAL, SEX) |>
  pivot_wider(names_from = PARAMCD, values_from = AVAL, values_fill = NA) 

dat_vif <- dat_wide |>
  select(-USUBJID)

formula <- as.formula(paste(c("~",paste(names(dat_vif), collapse="+")), collapse=""))
 
red <- Hmisc::redun(formula, data = dat_vif, pr=FALSE)
vif <- 1/(1-red$rsq1)
 
#cat("\nAvailable samvple size:\n", red$n, " (", #round(100*red$n/nrow(dat),2), "%)\n")

vif_df <- as_tibble(vif, rownames = "PARAMCD") |>
  rename(vif = value)

red_df <- as_tibble(red$rsq1, rownames = "PARAMCD") |>
  rename(r2 = value)

tab_df <- vif_df |> left_join(red_df, by = "PARAMCD")


```

The available sample size is `r red$n` (`r round(100*red$n/nrow(dat),2)` %).

```{r key_vif02, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}

#vif_df |> gt::gt(caption = "Variance inflation factors")
#red_df |> gt::gt(caption = "Multiple R-squared")

tab_df |> 
  gt() |>
    gt::cols_label(
    PARAMCD = "Parameter code",
    r2 = "Multiple R-squared",
    vif = "Variance inflation factor"
  ) |>
  fmt_number(
    columns = c(r2),
    decimals = 2
  ) |>
  fmt_number(
    columns = c(vif),
    decimals = 1
  ) |>
  gt_theme_538()
  

```


#### Redundancy by parametric additive model: key predictors and predictors of medium importance


```{r key_vif03, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}

dat_wide <- dat |>
  filter(KEY_PRED_FL02 == "Y" | MED_PRED_FL02 == "Y") |>
  select(USUBJID, PARAMCD, AVAL, SEX) |>
  pivot_wider(names_from = PARAMCD, values_from = AVAL, values_fill = NA) 


dat_vif <- dat_wide |>
  select(-USUBJID)

formula <- as.formula(paste(c("~",paste(names(dat_vif), collapse="+")), collapse=""))
 
#formula

red <- Hmisc::redun(formula, data = dat_vif, pr=FALSE)
vif <- 1/(1-red$rsq1)
 
#cat("\nAvailable samvple size:\n", red$n, " (", #round(100*red$n/nrow(dat),2), "%)\n")

vif_df <- as_tibble(vif, rownames = "PARAMCD") |>
  rename(vif = value)

red_df <- as_tibble(red$rsq1, rownames = "PARAMCD") |>
  rename(r2 = value)

tab_df <- vif_df |> left_join(red_df, by = "PARAMCD")


```

The available sample size is `r red$n` (`r round(100*red$n/nrow(dat),2)` %).

```{r key_vif04, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}

#vif_df |> gt::gt(caption = "Variance inflation factors")
#red_df |> gt::gt(caption = "Multiple R-squared")

tab_df |> 
  gt() |>
    gt::cols_label(
    PARAMCD = "Parameter code",
    r2 = "Multiple R-squared",
    vif = "Variance inflation factor"
  ) |>
  fmt_number(
    columns = c(r2),
    decimals = 2
  ) |>
  fmt_number(
    columns = c(vif),
    decimals = 1
  ) |>
  gt_theme_538()
  

```
