# Results of IDA: Missing values {#Missing}


```{r, warning=FALSE, message=FALSE}
## Load libraries
library(here)
library(dplyr)
library(tidyr)
library(naniar) ## for missing data 
library(gt)
library(gtExtras)
library(dendextend)

# Load the first iteration of the IDA ready data sets
ADSL <- readRDS(here::here("data", "IDA", "ADSL_01.rds"))
ADLB <- readRDS(here::here("data", "IDA", "ADLB_01.rds"))
```

THis seciton reports the IDA analyses for missing data. The Section headers (i.e. M2) correspond to the IDA analysis plan in @sec-IDA_plan.

## M1: Participant missingness

As the data was obtained from the registry of the laboratory, and only performed laboratory analyses were included, participant missingness cannot be evaluated.

## M2: Variable missingness

Number and percentage of missingness for each predictor, sorted by descending missingness proportion. 

### Outcome and Structural variables

```{r m1-structural, message = FALSE, warning = FALSE, echo = FALSE}

ADSL |>
  select(BACTEREMIA, AGE, SEX) |>
  miss_var_summary() |>
  gt() |>
    gt::cols_label(
    variable = "Variable",
    n_miss = "Missing (count)",
    pct_miss = "Missing (%)"
  ) |>
  fmt_number(
    columns = vars(pct_miss),
    decimals = 1
  ) |>
  tab_header(title = "Summary of missing values for outcome and structural predictors") |>
  gt_theme_538()

```

### Lab parameters 

Missingness patterns for key and medium importance predictors are reported here. The remaining predictors are reported in Appendix @sec-M1-appendix .

```{r m1-lab-predictors-key, message = FALSE, warning = FALSE, echo = FALSE}

source(here("R", "M1-predictor-missing.R"))

table_pred_missing_by_type(data = ABLB, flag = "KEY_PRED_FL01", tab_title = "Summary of missing values for key predictors")

```

```{r m1-lab-predictors-med, message = FALSE, warning = FALSE, echo = FALSE}

table_pred_missing_by_type(data = ABLB, flag = "MED_PRED_FL01", tab_title = "Summary of missing values for medium importance predictors")

```




## M3: Complete cases

Number of available complete cases (outcome and predictors):

```{r summiss, message =FALSE, warning =FALSE , echo=FALSE}

source(here("R", "M2-complete-cases.R"))
m2_ard <- m2_analysis(ADSL = ADSL, ADLB = ADLB)

m2_ard |> 
  select(Set, n_cases, pct_cases ) |>
  gt() |>
   cols_label(
    Set  = "Set",
    n_cases = "Complete (count)",
    pct_cases = "Complete (%)"
  ) |>
  fmt_number(
    columns = c(pct_cases),
    decimals = 1
  ) |>
  tab_header(title = "Summary of complete cases by sets of predictors") |>
  gt_theme_538() 

```


## M4: Patterns of missing values

### Complete cases by strata defined by structural variables

```{r summiss2, message =FALSE, warning =FALSE , echo=FALSE}

source(here("R", "M3-complete-cases-sets.R"))

m3_ard <- m3_analysis(ADSL, ADLB)

m3_ard |> 
  arrange(SEXC, AGEGR01) |>
  select(Set, n, n_cases, pct_cases, SEXC, AGEGR01C ) |>
  group_by(SEXC, AGEGR01C) |>
  gt() |>
   gt::cols_label(
    Set  = "Set",
    n = "# patients",
    n_cases = "Complete (count)",
    pct_cases = "Complete (%)"
  ) |>
  gt::fmt_number(
    columns = c(pct_cases),
    decimals = 1
  ) |>
  tab_header(title = "Summary of complete cases by sets of predictors") |>
  gt_theme_538()


```

### Dendrogram of missingness indicators 

The dendrogram depicts the results of a cluster analysis using the complete linkage method based on the percentage of discordant missing indicators. (This percentage was computed via the squared Euclidian distance of missingness indicators between predictors.) The horizontal axis shows the distance between two clusters, which is given by the maximum distance between any element of the first and the second clusters.  For example, if two clusters are merged at a height of 25 it means that in 25% of the observations the missingness indicators of the most discordant predictors contained in the two clusters are discordant.

The numbers in brackets are the percentages of missing observations for each predictor.

```{r m3-dendogram, message =FALSE, warning =FALSE , echo=FALSE, fig.height=13, fig.width=13}

#fig.cap="Clustered variables by percentage observations concordantly missing [by variable percentage missing]"

source(here("R", "M3-dendogram.R"))
m3_dendogram(ADLB)

```



