# Analysis ready data - second iteration {#ARD}


```{r ard01, warning=FALSE, message=FALSE, echo=FALSE}
## Load libraries for this chapter
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyselect)
library(here)
library(janitor)
library(gt)
library(gtExtras)
library(patchwork)
library(sjlabelled)


## Load the datasets - make sure to load the correct ADLB2 
## TODO: have a make workflow to ensure dependencies run in correct order

ADSL <- readRDS(here::here("data", "ADSL_01.rds"))
ADLB <- readRDS(here::here("data", "ADLB_02.rds"))

```

## dervive WBC_NEU

```{r}

source(here("R", "fun_ida_trans.R"))

wbc_no_neu_01 <-
  ADLB |>
  select(USUBJID, PARAMCD, AVAL) |>
  filter(PARAMCD %in% c("WBC", "NEU")) |>
  pivot_wider(names_from = "PARAMCD", values_from = "AVAL") |>
  mutate(AVAL = WBC - NEU,
         PARAMCD = "WBC_noNEU",
         PARAM = "White blood count without Neutrophiles (G/L)",
         PARAMTYPE = "DERVIVED") |>
  select(-WBC, -NEU)


wbc_no_neu_trans_01 <- ida_trans(wbc_no_neu_01)

wbc_no_neu_trans_02 <- wbc_no_neu_trans_01$dat

## plot side by side
source(here("R", "fun_compare_dist_plot.R"))

compare_dist_plot(wbc_no_neu_trans_02)

```

```{r}

ADLB_02 <- bind_rows(ADLB, wbc_no_neu_01, wbc_no_neu_trans_02) 
ADLB_02 <- sjlabelled::copy_labels(ADLB_02, ADLB)  
#ADLB_02 |> group_by(PARAMCD) |> tally()
```


# Set analysis flags for 


Handled by ADSL

  
Flags within ADBL

$pivotal_vars
[1] "WBC"   "AGE"

$vip_vars
[1] "WBC"   "AGE" "BUN"   "CREA"  "NEU"   "PLT"  

$leuko_related_vars
[1] "WBC"  "NEU"  "EOS"  "BASO" "LYM"  "MONO"

$leuko_ratio_vars
[1] "NEUR"  "EOSR"  "BASOR" "LYMR"  "MONOR"

$acute_related_vars
[1] "FIB"  "CRP"  "ASAT" "ALAT" "GGT" 

$remaining_vars
 [1] "MCV"  "HGB"  "HCT"  "MCH"  "MCHC" "RDW"  "MPV"  "NT"   "APTT" "NA."  "CA"   "PHOS" "MG"   "HS"   "GBIL"
[16] "TP"   "ALB"  "AMY"  "PAMY" "LIP"  "CHE"  "AP"   "LDH"  "CK"   "GLU"  "TRIG" "CHOL" "PDW"  "RBC" 




```{r set_flags, warning=FALSE, message=FALSE, echo=FALSE} 

ADLB02 <-
  ADLB |>
  mutate(
    KEY_PRED_FL03 = case_when(
      PARAMCD %in% c("AGE", "WBC_noNEU", "BUN", "CREA", "PLT", "NEU") ~ "Y"
      ),
    KEY_PRED_FL04 = case_when(
      PARAMCD %in% c("AGE", "WBC_noNEU_T", "BUN_T", "CREA_T", "PLT", "NEU_T") ~ "Y"
      )
    )


attr(ADLB02$KEY_PRED_FL03, "label") <- "Key Predictor Flag - original"
attr(ADLB02$KEY_PRED_FL04, "label") <- "Key Predictor Flag - transformed"


```

```{r}
vars <- ADLB02 |> 
  filter(KEY_PRED_FL03 == "Y") |> select(PARAMCD) |> distinct() |> as.list()
#paste(vars, collapse = ', ')

print(paste(vars, collapse = ', '))

```


`r print(paste(vars, collapse = ', '))`

# Add metadata flags to indicate relationship between blood cell parameters. See section 2.1.3 of the IDA plan 



## Complete cases flag

## key predictors orig

## key predictors transformed

