# Analysis ready data - second iteration {#ARD}


```{r ard01, warning=FALSE, message=FALSE, echo=FALSE}
## Load libraries for this chapter
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyselect)
library(here)
library(janitor)
library(gt)
library(gtExtras)
library(patchwork)



## Load the datasets - make sure to load the correct ADLB2 
## TODO: have a make workflow to ensure dependencies run in correct order

ADSL <- readRDS(here::here("data", "ADSL_01.rds"))
ADLB <- readRDS(here::here("data", "ADLB_02.rds"))

```

## dervive WBC_NEU

```{r}

source(here("R", "fun_ida_trans.R"))

wbc_no_neu_01 <-
  ADLB |>
  select(USUBJID, PARAMCD, AVAL) |>
  filter(PARAMCD %in% c("WBC", "NEU")) |>
  pivot_wider(names_from = "PARAMCD", values_from = "AVAL") |>
  mutate(AVAL = WBC - NEU,
         PARAMCD = "WBC_noNEU",
         PARAM = "WBC with NEU subtracted") |>
  select(-WBC, -NEU)


wbc_no_neu_trans_01 <- ida_trans(wbc_no_neu_01)

wbc_no_neu_trans_02 <- wbc_no_neu_trans_01$dat

## plot side by side
source(here("R", "fun_compare_dist_plot.R"))

compare_dist_plot(wbc_no_neu_trans_02)

```

```{r}

## TODO: check lineage of labels

ADLB_02 <- bind_rows(ADLB, wbc_no_neu_01, wbc_no_neu_trans_02) 
ADLB_02 <- sjlabelled::copy_labels(ADLB_02, ADLB)  

```


# Set analysis flags for 


Handled by ADSL

$demog_vars
[1] "Alter" "sex"  

$outcome_vars
[1] "BloodCulture" "BC" 

Flags within ADBL

$pivotal_vars
[1] "WBC"   "AGE"

$vip_vars
[1] "WBC"   "AGE" "BUN"   "CREA"  "NEU"   "PLT"  

$leuko_related_vars
[1] "WBC"  "NEU"  "EOS"  "BASO" "LYM"  "MONO"

$leuko_ratio_vars
[1] "NEUR"  "EOSR"  "BASOR" "LYMR"  "MONOR"

$acute_related_vars
[1] "FIB"  "CRP"  "ASAT" "ALAT" "GGT" 

$remaining_vars
 [1] "MCV"  "HGB"  "HCT"  "MCH"  "MCHC" "RDW"  "MPV"  "NT"   "APTT" "NA."  "CA"   "PHOS" "MG"   "HS"   "GBIL"
[16] "TP"   "ALB"  "AMY"  "PAMY" "LIP"  "CHE"  "AP"   "LDH"  "CK"   "GLU"  "TRIG" "CHOL" "PDW"  "RBC" 




```{r set_flags, warning=FALSE, message=FALSE, echo=FALSE} 
ADLB02 <-
  ADLB |>
  mutate(
    KEY_PRED_FL01 = case_when(
      PARAMCD %in% c("AGE", "WBC", "BUN", "CREA", "PLT", "NEU") ~ "Y"
      ),
    MED_PRED_FL01 = case_when(
      PARAMCD %in% c("POTASS", "FIB", "CRP", "ASAT", "ALAT", "GGT") ~ "Y"
      ),
    REM_PRED_FL01 = case_when(
      PARAMCD %in% c("MCV", "HGB", "HCT", "MCH", "MCHC", "RDW", "MPV", "LYM", "MONO", "EOS", "BASO", "NT", "APTT", "SODIUM", "CA", "PHOS", "MG", "HS", "GBIL", "TP", "ALB", "AMY", "PAMY", "LIP", "CHE", "AP", "LDH", "CK", "GLU", "TRIG", "CHOL", "BASOR", "EOSR", "LYMR", "MONOR", "NEUR", "PDW", "RBC") ~ "Y"
      )
  )


attr(bact_data04$KEY_PRED_FL01, "label") <- "Key Predictor Flag - Section 2.1.1 Analysis strategy"
attr(bact_data04$MED_PRED_FL01, "label") <- "Predictors of medium importance Flag - Section 2.1.1 Analysis strategy"
attr(bact_data04$REM_PRED_FL01, "label") <- "Remaining Predictors flag - Section 2.1.1 Analysis strategy"
```


Add metadata flags to indicate relationship between blood cell parameters. See section 2.1.3 of the IDA plan 



## Complete cases flag

## key predictors orig

## key predictors transformed

