# Source data {#sec-source-data}

```{r appendix-source-data-set-up, warning=FALSE, message=FALSE}
## Load libraries for this chapter
library(readr)
library(dplyr)
library(gt)
library(gtExtras)
library(janitor)
library(here)
```

The **source** data available in this repository (**DC 2019-0054**) can be located in the `data-raw` folder. For persistance, the data set is also published on [Zenodo](https://doi.org/10.5281/zenodo.7554815) with the following doi: https://doi.org/10.5281/zenodo.7554815. 

The following sections provide a short overview of the data dictionary which accompanies the source data, and a short description of the data. 


## Data dictionary {#sec-data_dict}

The data dictionary provides an overview of the collected data. First, we read and display the data dictionary below providing an overview of the collected measurements. 

```{r datadict, warning=FALSE, message=FALSE, echo=FALSE}

## Read in the source data dictionary, 
## clean column names to be consistent format 
## and arrange by variable order.  

bact_dd <- read_csv(
  here("data-raw", "Bacteremia_public_S2_Data_Dictionary.csv")
  ) |>
  arrange(VariableNr) |>
  clean_names()
```
The `variable` name and `label` are displayed alongside the measurement `scale` and `units` as well as `remarks` and relevant study information `from_paper`. 

```{r datadict_display}

## display the dictionary. 

bact_dd |> 
  select(-variable_nr, -remark, -from_paper) |>
  gt() |>
  gt_theme_538()

```

## Source data

We also display a short snapshot of source data set from the `data-raw` folder of the project directory. The snapshot provides a glimpse of the data, giving the data dictionary more context.  

```{r contents_bact,  warning=FALSE, message=FALSE, echo=FALSE} 

## read the source data set from the data folder.
bact <- read_csv(here("data-raw", "Bacteremia_public_S2.csv"))

```

We do not display all observations measured as it is too wide and long to fit reasonably in to the report. However, we refer you to the [Zenodo page](https://doi.org/10.5281/zenodo.7554815) for an interactive overview of the source data. 

```{r contents_bact_glimpse,  warning=FALSE, message=FALSE, echo=FALSE}

# TODO: Move to appendix?
bact |> glimpse(width = getOption("width"))

```

