# Univariate distribution checks {#sec-univar-appendix}

This section reports a series of univariate summary checks of the bacteremia dataset.

```{r uni01a, message =FALSE, warning =FALSE , echo=FALSE}
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyselect)
library(gt)
library(gtExtras)
library(patchwork)
library(Hmisc)
library(sjlabelled)

set.seed(1972)  ## for any simulation or numerical approx calculations 

# Load the first iteration of the IDA ready data sets
ADSL <- readRDS(here::here("data", "IDA", "ADSL_01.rds"))
ADLB <- readRDS(here::here("data", "IDA", "ADLB_01.rds"))
```

## U2: Descriptive summaries 

### U2: Remaining predictors {#sec-u2-remaining}

We present a visual summary. 

```{r uni04a, message=FALSE, warning=FALSE, echo=FALSE}
#| layout-ncol: 3
#| fig-width: 4
#| fig-height: 2

source(here::here("R", "U2-cont-describe-plot.R"))

rem_plts <-
  ADLB |>
  filter(REM_PRED_FL01 == "Y") |>
  group_by(PARAM) |>
  group_map(~ describe_plot(.x, num_bins = 300))

for (plts in rem_plts) {
  print(plts)
}

```

And a descriprive summary


```{r uni07, results='asis', message =FALSE, warning =FALSE , echo=FALSE}

remaining_predictors <-
  ADLB |>
  filter(REM_PRED_FL01 == "Y") |>
  select(USUBJID, PARAMCD, AVAL) |>
  pivot_wider(names_from = "PARAMCD", values_from = "AVAL") |>
  select(-USUBJID)


html(describe(remaining_predictors, size = "tiny"))

```

### Full descriptive summaries 

For U2, we present only a limited number of statistics are presented for brevity. However, a full set of descriptive summaries are available according to the specifications in the IDA plan as a data set. The summary statistics can be viewed and analysed in the following directory `data/results/U2-descriptive-stats.csv`. 

```{r full_desc, message = FALSE, warning = FALSE , echo = FALSE}
source(here::here("R", "fun_describe_long_dataset.R"))

# summary stats 
ARD <- describe_long_dataset(ADLB)

ARD |> pivot_wider(names_from = "STATISTIC", values_from = "RESULT") |>
  readr::write_csv(file = here("data", "results", "U2-descriptive-stats.csv"))

```


