# Pseudo-log transformations {.appendix}

```{r pl-setup, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This supplemental section illustrates how pseudo-log transformations can be used to transform skewed distributions towards normality. 

The transformation $f(x) = sinh^-1(x/2\sigma)/log(10)$ is a pseudo-logarithmic transformation mentioned by [@Johnson49]. It has the following advantages over ordinary logarithmic transformations:

* it is defined also for $x=$ as $f(0) = 0$
* it is a signed logarithmic transformation, and is defined also for negative values as $f(-|x|)=-f(|x|)$

Of course, this comes at the cost of deviation from the logarithmic transformation in terms of interpretability.

The parameter $\sigma$ may be used to adapt the transformation to a specific range of an empirical distribution.

We define the pseudo-logarithmic transformation in R as:

```{r pl02, message=FALSE, warning=FALSE, echo=TRUE, fig.height=3}
pseudo_log <-
  function(x, sigma = 1, base = 10)
    asinh(x / (2 * sigma)) / log(base)

inv_pseudo_log <-
  function(x, sigma = 1, base = 10)
    2 * sigma * sinh(x * log(base))
```

## Impact of choice of $\sigma$

Next, we investigate how the parameter $\sigma$ impacts the result of the transformation. We assume that $x$ follows a log-normal distribution, and we will show results of $f(x)$ with different choices of $\sigma$. We center and scale $f(x)$ before plotting.


```{r pl03, message=FALSE, warning=FALSE, echo=TRUE, fig.height=4}
p <- seq(0.001, 0.999, 0.001)
x <- exp(qnorm(p, mean = 0, sd = 1))

hist(x)

```



```{r pl03b, message=FALSE, warning=FALSE, echo=TRUE, fig.height=8}
y <-
  cbind(
    log(x),
    scale(pseudo_log(x, 1)),
    scale(pseudo_log(x, 0.5)),
    scale(pseudo_log(x, 2)),
    scale(pseudo_log(x, 0.05))
  )

plot(x,
     y[, 2],
     type = "l",
     ylab = "f(x)",
     ylim = range(y))
lines(x, y[, 1], col = "red")
lines(x, y[, 3], type = "l", lty = 2)
lines(x, y[, 4], type = "l", lty = 3)
lines(x, y[, 5], type = "l", lty = 4)
legend(
  "topleft",
  lty = c(1, 1, 2, 3, 4),
  col = c("red", "black", "black", "black", "black"),
  legend = c("log", "sigma=1", "sigma=0.5", "sigma=2", "sigma=0.05")
)
```


In the next code chunk, we multiply by 10 and repeat the exercise. We learn that the transformations become more similar and the choice of $\sigma$ less relevant.

```{r pl04, message=FALSE, warning=FALSE, echo=TRUE, fig.height=5}
x <- x * 10

y <-
  cbind(
    scale(log(x)),
    scale(pseudo_log(x, 1)),
    scale(pseudo_log(x, 0.5)),
    scale(pseudo_log(x, 2)),
    scale(pseudo_log(x, 0.05))
  )


plot(x, y[, 2], type = "l", ylab = "f(x)")
lines(x, y[, 1], col = "red")
lines(x, y[, 3], type = "l", lty = 2)
lines(x, y[, 4], type = "l", lty = 3)
lines(x, y[, 5], type = "l", lty = 4)
legend(
  "bottomright",
  lty = c(1, 1, 2, 3, 4),
  col = c("red", "black", "black", "black", "black"),
  legend = c("log", "sigma=1", "sigma=0.5", "sigma=2", "sigma=0.05")
)

```

Finally, we apply the pseudo-logarithmic transformation to the original normal deviates. We learn that a higher value for the parameter $\sigma$ makes the distribution 'slimmer' while a lower value makes it 'fatter', and it is even possible to induce bimodality with low values of sigma:

```{r pl05, message=FALSE, warning=FALSE, echo=TRUE, fig.height=5}
#| layout-ncol: 3 

z <- qnorm(p, mean = 0, sd = 1)
hist(z)
hist(pseudo_log(z, sigma = 1))
hist(pseudo_log(z, sigma = 2))
hist(pseudo_log(z, sigma = 0.5))
hist(pseudo_log(z, sigma = 0.05))
```

## Finding a parameter that best achieves normality

Any test statistic for testing normality could be chosen to find a suitable value of $\sigma$ that induces normality into the transformed values. Here we use the (Pearson) correlation coefficient to compare the empirical distribution with normal deviates.

### Deviates from log normal distribution

We simulate from a shifted log normal distribution and evaluate the value of $\sigma$ that optimizes agreement with a normal distribution:

```{r pl06, message=FALSE, warning=FALSE, echo=TRUE, fig.height=3}
#| layout-ncol: 2 

x <- sort(exp(rnorm(1000) + 3))

hist(x)

sigmas <- 2 ** seq(-10, 10, 1)
origcor <- cor(qnorm((1:length(x) - 0.5) / length(x)), x)

ncorx <-
  sapply(sigmas, function(X)
    cor(qnorm((1:length(
      x
    ) - 0.5) / length(x)), pseudo_log(x, X)))

cat("Optimal sigma: ")
(optsigma <- sigmas[ncorx == max(ncorx)])

plot(log2(sigmas), ncorx, ylab = "Correlation with normal deviates", ylim =
       c(0, 1))
points(log2(sigmas)[ncorx == max(ncorx)], max(ncorx), pch = 19)
abline(h = origcor)
legend(
  "bottomright",
  lty = c(1, NA, NA),
  pch = c(NA, 1, 19),
  legend = c("Original", "Pseudo-log", "Pseudo-log with optimal sigma")
)
box()

hist(pseudo_log(x, optsigma))
```

### Deviates from exponential distribution

Also with an exponential distribution, the pseudo-logarithm may achieve a better agreement with a normal:

```{r pl07, message=FALSE, warning=FALSE, echo=TRUE, fig.height=5}

x <- sort(exp(rnorm(1000) + 3))

hist(x)


sigmas <- 2 ** seq(-10, 10, 1)
origcor <- cor(qnorm((1:length(x) - 0.5) / length(x)), x)

ncorx <-
  sapply(sigmas, function(X)
    cor(qnorm((1:length(
      x
    ) - 0.5) / length(x)), pseudo_log(x, X)))

cat("Optimal sigma: ")
(optsigma <- sigmas[ncorx == max(ncorx)])

plot(log2(sigmas), ncorx, ylab = "Correlation with normal deviates", ylim =
       c(0, 1))
points(log2(sigmas)[ncorx == max(ncorx)], max(ncorx), pch = 19)
abline(h = origcor)
legend(
  "bottomright",
  lty = c(1, NA, NA),
  pch = c(NA, 1, 19),
  legend = c("Original", "Pseudo-log", "Pseudo-log with optimal sigma")
)
box()

hist(pseudo_log(x, optsigma))
```

### Deviates from mixture distribution

Now we mix a normal, lognormal and exponential distribution:

```{r pl08, message=FALSE, warning=FALSE, echo=TRUE, fig.height=5}

x1 <- scale(rnorm(1000))
x2 <- scale(rexp(1000))
x3 <- scale(exp(rnorm(1000)))

p1 <- rbinom(1000, 1, 0.33)
p2 <- rbinom(1000, 1, 0.5)

x <- p1 * x1 + (1 - p1) * (p2 * x2 + (1 - p2) * x3)
x <- x - min(x)
x <- sort(x)

hist(x)


sigmas <- 2 ** seq(-10, 10, 1)
origcor <- cor(qnorm((1:length(x) - 0.5) / length(x)), x)

ncorx <-
  sapply(sigmas, function(X)
    cor(qnorm((1:length(
      x
    ) - 0.5) / length(x)), pseudo_log(x, X)))

cat("Optimal sigma: ")
(optsigma <- sigmas[ncorx == max(ncorx)])

plot(log2(sigmas), ncorx, ylab = "Correlation with normal deviates", ylim =
       c(0, 1))
points(log2(sigmas)[ncorx == max(ncorx)], max(ncorx), pch = 19)
abline(h = origcor)
legend(
  "bottomright",
  lty = c(1, NA, NA),
  pch = c(NA, 1, 19),
  legend = c("Original", "Pseudo-log", "Pseudo-log with optimal sigma")
)
box()

hist(pseudo_log(x, optsigma))
```


### Normal deviates

With simulated normal deviates, the pseudo-logarithm cannot improve the already perfect normality.

```{r pl09, message=FALSE, warning=FALSE, echo=TRUE, fig.height=5}

x <- sort(rnorm(1000))

hist(x)

sigmas <- 2 ** seq(-10, 10, 1)
origcor <- cor(qnorm((1:length(x) - 0.5) / length(x)), x)

ncorx <-
  sapply(sigmas, function(X)
    cor(qnorm((1:length(
      x
    ) - 0.5) / length(x)), pseudo_log(x, X)))

cat("Optimal sigma: ")
(optsigma <- sigmas[ncorx == max(ncorx)])

plot(log2(sigmas), ncorx, ylab = "Correlation with normal deviates", ylim =
       c(0, 1))
points(log2(sigmas)[ncorx == max(ncorx)], max(ncorx), pch = 19)
legend(
  "bottomright",
  lty = c(1, NA, NA),
  pch = c(NA, 1, 19),
  legend = c("Original", "Pseudo-log", "Pseudo-log with optimal sigma")
)
abline(h = origcor)
box()

hist(pseudo_log(x, optsigma))
```

