# Source data {#sec-source-data}


```{r appendix-source-data-set-up, warning=FALSE, message=FALSE}
## Load libraries for this chapter
library(readr)
library(dplyr)
library(gt)
library(gtExtras)
library(janitor)
library(here)
```


## Data dictionary {#sec-data_dict}

The data dictionary provides an overview of the collected data. First, we read and display the data dictionary below providing an overview of the collected measurements. 

```{r datadict, warning=FALSE, message=FALSE, echo=FALSE}

## Read in the source data dictionary, 
## clean column names to be consistent format 
## and arrange by variable order.  

bact_dd <- read_csv(
  here("data-raw", "Bacteremia_public_S2_Data_Dictionary.csv")
  ) |>
  arrange(VariableNr) |>
  clean_names()
```

The `variable` name and `label` are displayed alongside the measurement `scale` and `units` as well as `remarks` and relevant study information `from_paper`. 

```{r datadict_display}

## display the dictionary. 

bact_dd |> 
  select(-variable_nr, -remark, -from_paper) |>
  gt() |>
  gt_theme_538()

```
